---
title: "Assessment 4 R Data Analysis"
author: "Natasha Ford, Kahla Finch, Devanshee Alpeshkumar Gandhi"
date: "2024-09-25"
output: html_document
fig_width: 10
---

# **Part 1**


## Download Data files

```{r Download Data}
URL1="https://raw.githubusercontent.com/ghazkha/Assessment4/refs/heads/main/growth_data.csv"
download.file(URL1,destfile="Growth_Data.csv")

URL2="https://raw.githubusercontent.com/ghazkha/Assessment4/refs/heads/main/gene_expression.tsv"
download.file(URL2,destfile="Gene_Expression.csv")
```


## Gene Expression Data
The file “gene_expression.tsv” contains RNA-seq count data for three samples of interest.


### Read in the file, making the gene identifiers the row names. Show a table of values for the first six genes.

```{r Read in File}
library(readr)
Gene_Expression <- read.table("https://raw.githubusercontent.com/ghazkha/Assessment4/refs/heads/main/gene_expression.tsv",stringsAsFactors = FALSE, header=TRUE, row.names=1)

head(Gene_Expression)
```


### Make a new column which is the mean of the other columns. Show a table of values for the first six genes.
```{r Calculating Mean values}
Gene_Expression$Mean_value <- rowMeans(Gene_Expression, na.rm = TRUE)
head(Gene_Expression)
```


### List the 10 genes with the highest mean expression
```{r 10 genes with the highest mean expression}
Sorted_Genes <- Gene_Expression[order(-Gene_Expression$Mean_value), ]
Top10_Mean <- head(Sorted_Genes, 10)
print(Top10_Mean)
```

### Determine the number of genes with a mean <10
```{r Number of genes with mean less than 10}

Low_Mean <- subset(Gene_Expression, Mean_value < 10)
nrow(Low_Mean)
```


### Session Info
```{r}
sessionInfo()
```

