---
title: "Assessment 4 R Data Analysis"
author: "Natasha Ford, Kahla Finch, Devanshee Alpeshkumar Gandhi"
date: "2024-09-25"
output: html_document
fig_width: 10
---

# **Part 1**


## Download Data files

```{r Download Data}

URL2="https://raw.githubusercontent.com/ghazkha/Assessment4/refs/heads/main/gene_expression.tsv"
download.file(URL2,destfile="Gene_Expression.csv")
```


## Gene Expression Data
The file “gene_expression.tsv” contains RNA-seq count data for three samples of interest.


### Read in the file, making the gene identifiers the row names. Show a table of values for the first six genes.

```{r Read in File}
install.packages("readr")
library(readr)
Gene_Expression <- read.table("https://raw.githubusercontent.com/ghazkha/Assessment4/refs/heads/main/gene_expression.tsv",stringsAsFactors = FALSE, header=TRUE, row.names=1)

head(Gene_Expression)
```


### Make a new column which is the mean of the other columns. Show a table of values for the first six genes.
```{r Calculating Mean values}
Gene_Expression$Mean_value <- rowMeans(Gene_Expression, na.rm = TRUE)
head(Gene_Expression)
```


### List the 10 genes with the highest mean expression
```{r 10 genes with the highest mean expression}
Sorted_Genes <- Gene_Expression[order(-Gene_Expression$Mean_value), ]
Top10_Mean <- head(Sorted_Genes, 10)
print(Top10_Mean)
```

### Determine the number of genes with a mean <10
```{r Number of genes with mean less than 10}

Low_Mean <- subset(Gene_Expression, Mean_value < 10)
nrow(Low_Mean)
```

### Histogram of Mean Values
```{r}
hist(Gene_Expression$Mean_value,breaks=1000, xlim = c(0,60000), xlab = "Mean Value of Gene Expression",main="Histogram of Mean values of Gene Expression")

```

The Histogram shown in the above figure displays the mean values of the gene expression data.

### Importing growth data csv file into R
```{r}
growth_data_URL <- "https://raw.githubusercontent.com/ghazkha/Assessment4/refs/heads/main/growth_data.csv"
download.file(growth_data_URL,destfile="growth_data")

growth_data <- read.csv("growth_data")
head(growth_data)
```

### Column names of Growth Data 
```{r}
colnames(growth_data)
```

### Mean and standard deviation at notheast and southwest sites at 2005 and 2020
####Mean and standard deviation of northeast site at 2005
```{r}
growth_northeast_2005 <- subset(growth_data, Site == "northeast", select = c("Circumf_2005_cm"))
growth_northeast_2005 

mean_growth_northeast_2005 <- mean(growth_northeast_2005$Circumf_2005_cm)
mean_growth_northeast_2005

SD_growth_northeast_2005 <- sd(growth_northeast_2005$Circumf_2005_cm)
SD_growth_northeast_2005
```

####Mean and standard deviation of northeast site at 2020
```{r}
growth_northeast_2020 <- subset(growth_data, Site == "northeast", select = c("Circumf_2020_cm"))
growth_northeast_2020

mean_growth_northeast_2020 <- mean(growth_northeast_2020$Circumf_2020_cm)
mean_growth_northeast_2020

SD_growth_northeast_2020 <- sd(growth_northeast_2020$Circumf_2020_cm)
SD_growth_northeast_2020
```


####Mean and standard deviation of southwest site at 2005
```{r}
growth_southwest_2005 <- subset(growth_data, Site == "southwest", select = c("Circumf_2005_cm"))
growth_southwest_2005

mean_growth_southwest_2005 <- mean(growth_southwest_2005$Circumf_2005_cm)
mean_growth_southwest_2005

SD_growth_southwest_2005 <- sd(growth_southwest_2005$Circumf_2005_cm)
SD_growth_southwest_2005
```

####Mean and standard deviation of southwest site at 2020
```{r}

growth_southwest_2020 <- subset(growth_data, Site == "southwest", select = c("Circumf_2020_cm"))
growth_southwest_2020

mean_growth_southwest_2020 <- mean(growth_southwest_2020$Circumf_2020_cm)
mean_growth_southwest_2020

SD_growth_southwest_2020 <- sd(growth_southwest_2020$Circumf_2020_cm)
SD_growth_southwest_2020
```











### Session Info
```{r}
sessionInfo()
```

